### 2013-01-05 sections/challenges.tex
moved sections into a separate dir

%!TEX root = paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{sec:introduction}
%%%%%%%%%%%%%%%%%%%%%%