### 2013-01-05 sections/implications.tex

%!TEX root = paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{Recommendations and Implications} \label{sec:implications}
%%%%%%%%%%%%%%%%%%%%%%