### 2013-01-06 sections/challenges.tex
done up to section IV (partially included)

%!TEX root = ../paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{What Are The Challenges of Code Review?} \label{sec:challenges}
%%%%%%%%%%%%%%%%%%%%%%