### 2013-01-06 sections/conclusion.tex

%!TEX root = ../paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion} \label{sec:conclusion}
%%%%%%%%%%%%%%%%%%%%%%