### 2013-01-06 sections/implications.tex

%!TEX root = ../paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{Recommendations and Implications} \label{sec:implications}
%%%%%%%%%%%%%%%%%%%%%%