### 2013-01-06 sections/limitations.tex

%!TEX root = ../paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{Limitations} \label{sec:limitations}
%%%%%%%%%%%%%%%%%%%%%%