### 2013-01-06 sections/outcomes.tex

%!TEX root = ../paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{The Outcomes of Code Reviews} \label{sec:outcomes}
%%%%%%%%%%%%%%%%%%%%%%