### 2013-01-06 sections/conclusion.tex
done with section 8 and 9

%!TEX root = ../paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion} \label{sec:conclusion}
%%%%%%%%%%%%%%%%%%%%%%

In this work, we have investigated modern, tool-based code review, uncovered both a wide range of motivations for review, and determined that the outcomes do not always match those motivations. We identified understanding as a key component and provided recommendations to both practitioners and researchers. It is our hope that the insights we have discovered lead to more effective review in practice and improved tools, based on research, to aid developers perform code reviews.