### 2013-03-01 sections/conclusion.tex
did another full pass on the paper

%!TEX root = ../paper.tex

%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion} \label{sec:conclusion}
%%%%%%%%%%%%%%%%%%%%%%

We investigated modern, tool-based code review, uncovering both a wide range of motivations for review and that the outcomes do not always match those motivations. We identified understanding as a key component and provided recommendations to both practitioners and researchers. It is our hope that the insights we have discovered lead to more effective review in practice and improved tools, based on research, to aid developers perform code reviews.