### 2013-03-16 paper.tex
changed according to camera-ready formatting instructions

\documentclass[conference]{IEEEtran}

\usepackage{graphicx}\graphicspath{{figs/}}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{listings}
\usepackage{ifthen}
\usepackage{float} % defines newfloat used below
%\usepackage[sort&compress]{natbib} % <<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<< commented out
%\usepackage{todos} % Use \todoset{all=off} to turn off todos in a paper.
\usepackage{booktabs}
\usepackage{fancyhdr}
\usepackage{microtype}
\usepackage{paralist}

% hyperref redefines a number of macros, so it should be last.  Empirically,
% doing so eliminates compiler warnings.
\usepackage{hyperref}
% According to the hyperref readme, algorithm must follow hyperref

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Macros

% English
\newcommand{\cf}{\hbox{\emph{cf.}}\xspace}
\newcommand{\deletia}{\ldots [deletia] \ldots}
\newcommand{\etal}{\hbox{\emph{et al.}}\xspace}
\newcommand{\eg}{\hbox{\emph{e.g.},}\xspace}
\newcommand{\ie}{\hbox{\emph{i.e.},}\xspace}
\newcommand{\scil}{\hbox{\emph{sc.}}\xspace} %scilicet: it is permitted to know
\newcommand{\st}{\hbox{\emph{s.t.}}\xspace}
\newcommand{\wrt}{\hbox{\emph{w.r.t.}}\xspace}
\newcommand{\etc}{\hbox{\emph{etc.}}\xspace}
\newcommand{\viz}{\hbox{\emph{viz.}}\xspace} %videlicet: it is permitted to see
\newcommand{\vs}{\hbox{\emph{vs.}}\xspace}

\newcommand{\cat}[1]{{\textsc{#1}}}

\usepackage{paralist}
\usepackage{mathrsfs}
\usepackage{tabularx}
\usepackage{subfigure}

\newcommand{\ab}[1]{\todo[Alberto]{#1}}
\newcommand{\cb}[1]{\todo[Chris]{#1}}

\renewcommand{\quotation}[1]{``\emph{#1}''}

\iffalse{
\def\todo#1{{\sf $\natural$ #1 $\natural$}}
\def\note#1{{\sf $\clubsuit$ #1 $\clubsuit$}}
\def\marknote#1{{\sf $\spadesuit$ #1 $\spadesuit$}}
}\fi

%\usepackage{mdframed} <<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<< commented out
%% Document-specific hyperref options
\hypersetup{
pdftitle={Expectations, Outcomes, and Challenges of Modern Code Review},
    pdfauthor={Alberto Bacchelli and Christian Bird},
    plainpages=false,
    linkcolor=black, % Overriding these colors to black is somewhat unfortunate 
    %citecolor=black, % b/c the defaults are useful in color.
    citecolor=blue, % b/c the defaults are useful in color.
    filecolor=black,
    urlcolor=black,
    pdfpagelabels
}

%% sig-alternative is stupid
%\pdfpagewidth=8.5in
%\pdfpageheight=11in

% for our nicer-looking \paragraphs under sig-alternate
\newlength{\emstr}
\setlength{\emstr}{0.75em plus 1ex minus 1ex}
\newcommand{\boldpara}[1]{%
\vspace{1.5mm}%
\par\noindent\textbf{#1}\hspace{\emstr}
}%

\newcommand{\figref}[1]{Figure~\ref{#1}\xspace}

\newcounter{copyrightbox}

\title{Expectations, Outcomes, and Challenges \\ 
of Modern Code Review}

\author{
	\begin{tabular}{ccc}
		Alberto Bacchelli & \hspace{0.5in} & Christian Bird \\
		REVEAL @ Faculty of Informatics & & Microsoft Research \\
		University of Lugano, Switzerland & & Redmond, Washington, USA \\
		alberto.bacchelli@usi.ch &  & cbird@microsoft.com
	\end{tabular}
}

%\iffalse{
% Only for camera ready -- prevent TeX from rendering widows
%\widowpenalty=9999
%\clubpenalty=9999 
%}\fi

\begin{document}

\maketitle

\begin{abstract}

Code review is a common software engineering practice employed both in open
source and industrial contexts.  Review today is less formal and more
``lightweight'' than the code inspections performed and studied in the 70s and
80s. We empirically explore the motivations, challenges, and outcomes of
tool-based code reviews. We observed, interviewed, and surveyed developers and
managers and manually classified hundreds of review comments across diverse
teams at Microsoft. Our study reveals that while finding defects remains the
main motivation for review, reviews are less about defects than expected and
instead provide additional benefits such as knowledge transfer, increased team
awareness, and creation of alternative solutions to problems. Moreover, we find
that code and change understanding is the key aspect of code reviewing and that
developers employ a wide range of mechanisms to meet their understanding needs,
most of which are not met by current tools. We provide recommendations for
practitioners and researchers.

\end{abstract}

%\listoftodos

\input{sections/introduction}
\input{sections/related_work}
\input{sections/methodology}
\input{sections/expectations}
\input{sections/outcomes}
\input{sections/challenges}
\input{sections/implications}
\input{sections/limitations}
\input{sections/conclusion}

\bibliographystyle{abbrv}
%\bibliographystyle{plain}
%\setlength{\bibsep}{3pt}
\small
\bibliography{paper}

\end{document}